Introduction to Structural Causal Models

### 2021-05-21 presentation_video.Rmd
more

# Fundamental Differences

- Conflict whether to use **graphs** or not and to use **parameters** in SEM
* SEM is a parametric model used in applied sciences 
* A Bayesian causal network (BCN) use conditional probabilities instead of functions 
* PO is mostly equivalent but has not graphical framework
* SCM and PO are only frameworks that allow for counterfactuals
- Every SCM has the same information as respective probabilistic model + more (DGP)
- SCMs can derive potential outcomes but PO cannot derive SCMs

@@

- Association-based learning is easy to model because of lower information neccessary
- but not always appropriate in high stake settings 
- SCM as simulator for causal modelling based on noise variables
- SCMS entail a lot of information (DGP, intervention distribution)
- Note: *"Garbage in, Garbage out"*
- computational advantage casting causal model as assignments 
- Enables modelling of higher order concepts like counterfactuals 
- Extensions through differential equations for concise modelling of time

@@

* ...only have conditional probabilities 
* ...are only interested in association 
* ...have very limited knowledge of your actual underlying structure

@@

* ... have an understanding of underlying relationships in the data
* ... are dealing with high stake settings 
* ... want to simulate different interventions

$\rightarrow$ you should probably go for a SCM

>>>>

# Differences

**SEM:**

- Conflict whether to use **graphs** or not and to use **parameters** in SEM
- SEM is a parametric model used in applied sciences

**BCN**:

- A Bayesian causal network (BCN) use conditional probabilities instead of functions 
- Every SCM has the same information as respective probabilistic model + more (DGP)
- Computationally less efficient and cannot cast counterfactuals

**Potential Outcome Framework:**

- PO is mostly equivalent but has not graphical framework
- SCMs can derive potential outcomes but PO cannot derive SCMs

@@

- Association-based learning is easy to model because of lower information neccessary
- but not always appropriate in high stake settings 
- SCM as simulator for causal modelling based on noise variables
- SCMs entail a lot of information (DGP, intervention distribution)
- BUT: **"Garbage in, Garbage out"**
- Computational advantage casting causal model as assignments 
- Enables modelling of higher order concepts like counterfactuals 
- Extensions through differential equations for concise modelling of time

@@

* ..only have conditional probabilities 
* ..are only interested in association 
* ..have very limited knowledge of your actual underlying structure

@@

* ..have an understanding of underlying relationships in the data
* ..are dealing with high stake settings 
* ..want to simulate different interventions

$\rightarrow$ you could probably go for a SCM
